\vspace*{\fill}
\begin{abstract}
In this bachelor thesis we present a 3D face registration pipeline by adapting
Gaussian Process Regression to use labelled point and line correspondence
as prior information. Gaussian Processes are ideal for modelling
smooth deformations of an arbitrary shape surface. Therefore, Gaussian
Process Regression can be implemented to sample deformation fields
that map a template on to a target shape using a set of corresponding
facial landmarks. In our approach we enhance the registration of face scans
by incorporating additional information from line features. These line features
define correspondence in highly characteristic face areas, i.e.
the eyes and the mouth. The deformation fields are optimized - using the
prior information as constraints - to obtain a plausible mapping of the
template on to the target. We demonstrate that line features have a profound
impact on the accuracy and quality of the registration results. In order to
enhance the expressiveness of the faces obtained through registration, we plan to
incorporate additional knowledge tailored to 3D face registration into
Gaussian Process Regression and further tune the optimization process.
\begin{comment}
In this bachelor thesis we present a 3D face registration pipeline by adapting Gaussian Process Regression and using labelled point and line correspondences as prior information for the registration process.
Gaussian Processes are ideal for modelling smooth deformations of an arbitrary shape surface. 
Therefore, Gaussian Process Regression can be implemented to sample deformation fields that map a template on to a target shape using a set of corresponding facial landmarks. In our approach we extend the registration of face scans by incorporating additional information from line features. The line features define correspondence in highly characteristic areas of a set of faces, i.e. the eyes and the ears. The deformation fields are optimized - using the prior information as
constraints - to
obtain a plausible mapping of the template on to the target. We demonstrate that line features have a profound impact on the accuracy and quality of the registration results. In order to enhance the expressiveness of the obtained registrations, we plan to further adjust the parameters of Gaussian Process Regression to the case of 3D faces and to tweak the optimization process in the near future. 

In this bachelor thesis we present the construction of a 3D face registration pipeline in a two-facetted approach. Firstly, we derive the registration algorithm from Gaussian Process Regression. Secondly, we augment registration quality by passing line correspondences to the algorithm as prior information.

In order to use this approach for model building, however, optimization of the expressiveness of the deformed template is still necessary.

 by using an algorithm based on a vector-valued gaussian process and at the same time attempting to ensure registration quality through the use of parametric curves - referred to as line features - that describe complex regions of the face.
from parametric curves - referred to as line features - that define key features of the faces, 


In this bachelor thesis we attempt to modify the existing face registration pipeline for the morphable face model of Prof. Thomas Vetter by using a registration algorithm developed by PD Marcel Lüthi at the University of Basel. ALTERNATIVE:
\end{comment}
\begin{comment}
The algorithm, which is derived from Gaussian Process Regression is capable to map any two shapes on to one another without much modification. It uses prior information in the form of a set of corresponding points on the two shapes in order to predic 
The algorithm is capable of mapping any two shapes on to one another. All that is needed is a set of corresponding points on the two shapes. Different constraints to the displacement field can be applied through regularisation. 

The aim of this bachelor thesis is more specifically to apply this general algorithm for point correspondences to scanned face data, that is to implement feasible registration of face scans onto the mean face of the morphable model. 
In order to achieve this we mark important parts of the face meshs not only with point landmarks, but also structures and organs (eyebrows, eyes, ears) with lines - line features - and thereby to create further correspondences for the algorithm to perform better by. 
Instead of using sparse points of key features points of the face we mark complex features, e.g. the eyes, with contour lines - line features in order to create further correspondences 

These line features are marked by hand using bézier curves on three 2D images to the front, left and right of the 3D face. 
In order to utilize them, however, they have to be projected on to the computed mesh of the face that was recorded by a 3D scanner. These meshs have holes in the region of the eyes and the ears rendering the projected line features useless at first. This thesis first gives an overview over the morphable model and the face registration pipeline, then goes on to obtaining 3D points from the 2D line features, to explain the theory behind the general algorithm and in the main part discusses
the problems and solutions we encountered trying to optimize the algorithm for and without line features for the face registration process.
\end{comment}
\end{abstract}
\vspace*{\fill}
\pagebreak
